---
title: 'Lab week 5: Time series and a map'
author: "Shellby Johnson"
date: "February 6, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```



```{r}

library(lubridate)
library(janitor)
library(tidyverse)
library(paletteer)
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forecast)
library(sf)
library(tmap)
library(here)
library(mapview)

```



##Monthly US energy consumption (renwables)

```{r}
us_renew <- read_csv(here("data", "renewables_cons_prod.csv")) %>% 
  clean_names()

```


- Make the description all lowercase and keep obervations where the description variable comtains the string pattern "consumption""
-Also gonna remove any observation where the "description"" variable contains "total"

```{r}
renew_clean <- us_renew %>% 
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, pattern = "consumption")) %>% 
  filter(!str_detect(description, pattern = "total")) #! means do not detect, so if it runs across an observation that does not have "consumption, it will keep it"

```

### Convert 'yyyymm' column to a date

```{r}

renew_date <- renew_clean %>% 
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm,"ym")) %>% 
  mutate(month_sep = yearmonth(yr_mo_day)) %>%# makes the date column only a month abbreviation and a year and separates the day
  mutate(value = as.numeric(value)) %>% 
  drop_na(month_sep, value)

#Make a version where teh month and year is in separate columns

renew_parse <- renew_date %>% 
  mutate(month = month(yr_mo_day, label = TRUE)) %>% 
  mutate(year = year(yr_mo_day))## creates a column of a class of ordered factors that only contains the month abbreviation

```



##Look at it.

```{r}

renew_gg <- ggplot(data = renew_date, aes(x = month_sep,
                                          y = value,
                                          group = description))+
                                          geom_line(aes(color = description))+
                                          theme_classic()+
                                          scale_y_continuous(0,350)

renew_gg
                   
                

```


##Updating my colors with paletter palettes

```{r}

renew_gg +
  scale_color_paletteer_d("calecopal::figmtn")

```


### Coerce our renew_parse to a sibble, a time-series enables data frame

```{r}

renew_ts <- as_tsibble(renew_parse, key = description, index = month_sep)

##index is the sibble compatible time variable that we created (the 'month_sep)

```

Lets look at our time series data in a couple different ways
```{r}

renew_ts %>% autoplot(value)
renew_ts %>% gg_subseries(value)
##renew_ts %>% gg_season(value) did not work

ggplot(data = renew_parse, aes(x = month, y = value, group = year))+
  geom_line(aes(color = year))+
  facet_wrap(~description,
             ncol = 1,
             scales = "free", 
             strip.position = "right")

```


##Just look at the hydroelectric eneergy consumption

```{r}

hydro_ts <- renew_ts %>% 
  filter(description == "hydroelectric power consumption")

##hydro_ts %>% autoplot(value)

ggplot(hydro_ts, aes(x = month, y = value, group = year))+
  geom_line(aes(color = year))

```


### What is I want quarterly average consumption for hydro?

```{r}

hydro_quarterly <- hydro_ts %>% 
  index_by(year_qu = ~ yearquarter(.)) %>% 
  summarize(avg_consumption = mean(value))

head(hydro_quarterly)

```

### Decompose that hydro_ts

```{r}
dcmp <- hydro_ts %>% 
  model(STL(value ~ season(window = 5)))


components(dcmp) %>% autoplot()


```


```{r}
hist(components(dcmp)$remainder)

```

# Now lets look at the ACF

```{r}

 hydro_ts %>% 
  ACF(value) %>% 
  autoplot()

## ACF shows that observations that are 12 months apart are more correlated than other time period
```

# DANGER DANGER

```{r}
hydro_model <- hydro_ts %>% 
  model(
    ARIMA(value)
  ) %>% 
fabletools::forecast(h = "4 years")

hydro_model %>% autoplot(filter(hydro_ts, year(month_sep) > 2010))

```

### Now lets make a world map

```{r}

world <- read_sf(dsn = here("data", "TM_WORLD_BORDERS_SIMPL-0.3-1"),
                 layer = "TM_WORLD_BORDERS_SIMPL-0.3")

mapview(world)

```

